The error in Bayesian phylogenetic reconstruction when speciation is not instantaneous

### 2018-06-14 article.tex
Rates can differ

The PBD model has five biological parameters, depicted in table \ref{table:parameters}, 
which we explore in a factorial fashion, excluding - for computational reasons - the combinations in which the 
95\% quantile of the expected number of good species is more than 1250.
This quantile is
calculated with the \verb;pbd_numspec_quantile; function we added to the \verb;PBD; package (\cite{pbd}).
\RAMPAL{I think we should release the new version with this function with this manuscript}\richel{I would personally prefer 'release early, release often', but as lead maintainer you get to decide}\RAMPAL{Do you want to put the derivation here?}\richel{Yes please, either here or at the PBD package (the functions pbd\_geom and 
pbd\_numspec\_quantile) documentation. Wherever you put it, I will transfer it to the other place}. This calculation
assumes $b = b_g = b_i$, we used $b = \max(b_g, b_i)$.
We use 1000 good species as a threshold, to prevent overly taxon-poor and taxon-rich phylogenies respectively.
The parameter values chosen are based on the parameter sets used by 
\cite{etienne2014}, as these parameters were shown to result in reasonably
sized phylogenies and using the same set allows us to compare results. 
We use a set of speciation initiation rates, $B = \{0.3, 0.5\}$, of which the speciation initiation
rate of good species $b_g \in B$ and incipient species $b_i \in B$. \richel{these rates can differ, although we do not plot the combinations in which they differ}\RAMPAL{Prima om die verschillend te nemen. Het ms was hier idd niet duidelijk over.}.\RAMPAL{your units are not entirely correct, as these are probability rates.}\richel{I though I was correct in my units. If I am wrong, what are the correct units then? In Etienne and Rosindell 2012, there is never 'probability rate' written. To be explicit, I think these sentences are all valid: (1) the extinction rate of good species, mu\_g is 1.2 extinction events per time unit (2) the speciation initiation rate of good species, b\_g is 2.3 speciation events per time unit, (3) the speciation completion rate, lambda is 3.4 speciation completion events per time unit. What would be the better unit?}\RAMPAL{
Je hebt het over de eenheden van de rates. Dat is subtiel omdat het om een stochastisch model gaat en dus om kansen. Voorbeeld: mu * dt is de kans op een extinction event per species in een tijdsinterval dt. 1/mu geeft de verwachte tijd tot een extinctie event van een soort. mu is dus een probability rate.
}
The speciation completion rates we use are $\lambda = 0.1$, $0.3$, $1.0$ and $10^9$. We use $10^9 \approx \infty$
to mimic the BD model, because the PBD model reduces to the BD model for $\lambda = \infty$.
This allows us to measure the baseline error, 
which is the difference between inferred tree and true species tree that arises purely due to noise 
because the generating model and the model used in inference are identical in this case.
We use a set of extinction rates, $M = \{0.0, 0.1, 0.2, \infty\}$, of which the extinction
rate of good species $\mu_g \in M$ and incipient species $\mu_i \in M$. 
\richel{Also here: these rates can differ}

>>>>

The PBD model has five biological parameters, depicted in table \ref{table:parameters}, 
which we explore in a factorial fashion, excluding - for computational reasons - the combinations in which the 
95\% quantile of the expected number of good species is more than 1250.
This quantile is
calculated with the \verb;pbd_numspec_quantile; function we added to the \verb;PBD; package (\cite{pbd}).
\RAMPAL{I think we should release the new version with this function with this manuscript}\richel{I would personally prefer 'release early, release often', but as lead maintainer you get to decide}\RAMPAL{Do you want to put the derivation here?}\richel{Yes please, either here or at the PBD package (the functions pbd\_geom and 
pbd\_numspec\_quantile) documentation. Wherever you put it, I will transfer it to the other place}. This calculation
assumes $b = b_g = b_i$, we used $b = \max(b_g, b_i)$.
We use 1000 good species as a threshold, to prevent overly taxon-poor and taxon-rich phylogenies respectively.
The parameter values chosen are based on the parameter sets used by 
\cite{etienne2014}, as these parameters were shown to result in reasonably
sized phylogenies and using the same set allows us to compare results. 
We use a set of speciation initiation rates, $B = \{0.3, 0.5\}$, of which the speciation initiation
rate of good species $b_g \in B$ and incipient species $b_i \in B$.\RAMPAL{your units are not entirely correct, as these are probability rates.}\richel{I though I was correct in my units. If I am wrong, what are the correct units then? In Etienne and Rosindell 2012, there is never 'probability rate' written. To be explicit, I think these sentences are all valid: (1) the extinction rate of good species, mu\_g is 1.2 extinction events per time unit (2) the speciation initiation rate of good species, b\_g is 2.3 speciation events per time unit, (3) the speciation completion rate, lambda is 3.4 speciation completion events per time unit. What would be the better unit?}\RAMPAL{
Je hebt het over de eenheden van de rates. Dat is subtiel omdat het om een stochastisch model gaat en dus om kansen. Voorbeeld: mu * dt is de kans op een extinction event per species in een tijdsinterval dt. 1/mu geeft de verwachte tijd tot een extinctie event van een soort. mu is dus een probability rate.
}
The speciation completion rates we use are $\lambda = 0.1$, $0.3$, $1.0$ and $10^9$. We use $10^9 \approx \infty$
to mimic the BD model, because the PBD model reduces to the BD model for $\lambda = \infty$.
This allows us to measure the baseline error, 
which is the difference between inferred tree and true species tree that arises purely due to noise 
because the generating model and the model used in inference are identical in this case.
We use a set of extinction rates, $M = \{0.0, 0.1, 0.2, \infty\}$, of which the extinction
rate of good species $\mu_g \in M$ and incipient species $\mu_i \in M$.
